---
title: "DS501 Case Study 3"
author: "Louis Duh"
date: "11/30/2021"
output: pdf_document
---

##Introduction to Case Study

For this case study, I will be creating an R shiny application as a data product based on a dataset. I am quite interested in NFL football and I also find oddsmaking and topics around how sportsbetting works to be interesting. Through browsing datasets, I determined that my topic for this data product will be NFL scores and betting data. I found a dataset on kaggle that includes data on these two topics and I will be using that to create my application. This dataset includes infomration such as stadium, whether it is indoor or outdoor, weather, week of the season, favroites, spread, and more. I will utilize this data set to analyze and show relations between any of these conditions, and the outcome of the game for both NFL and betting purposes. The goal is to have an application that can assist in predicting the outcome of NFL games based on the data set available. 

##Create and Format Dataframe
```{r, echo=TRUE, eval=TRUE}
#Library setup
library(dplyr)
library (tibble)
library(plotly)

#Import data file and Read
NFL_Data = read.csv("spreadspoke_scores.csv")
NFL_StadiumInfo = read.csv("nfl_stadiums.csv")
NFL_TeamInfo = read.csv("nfl_teams.csv")

#Create Column for total points in game
NFL_Data = NFL_Data %>% add_column(Total_Points = NFL_Data$score_home + NFL_Data$score_away, .after = "team_away")

#Create Column for margin of victory
NFL_Data = NFL_Data %>% 
  mutate(Victory_Margin = if_else(score_home>score_away,score_home-score_away,score_away-score_home))

#Limit to Data of seasons 2000 and later
NFL_Data_After2000 = subset(NFL_Data, schedule_season >= 2000)

#Replace Home Team column with Home Team ID 
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Arizona Cardinals"] = "ARI"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Atlanta Falcons"] = "ATL"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Baltimore Ravens"] = "BAL"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Buffalo Bills"] = "BUF"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Carolina Panthers"] = "CAR"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Chicago Bears"] = "CHI"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Cincinnati Bengals"] = "CIN"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Cleveland Browns"] = "CLE"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Dallas Cowboys"] = "DAL"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Denver Broncos"] = "DEN"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Detroit Lions"] = "DET"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Green Bay Packers"] = "GB"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Houston Texans"] = "HOU"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Indianapolis Colts"] = "IND"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Jacksonville Jaguars"] = "JAX"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Kansas City Chiefs"] = "KC"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Los Angeles Chargers"] = "LAC"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "San Diego Chargers"] = "LAC"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Los Angeles Rams"] = "LAR"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "St. Louis Rams"] = "LAR"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Miami Dolphins"] = "MIA"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Minnesota Vikings"] = "MIN"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "New England Patriots"] = "NE"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "New Orleans Saints"] = "NO"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "New York Giants"] = "NYG"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "New York Jets"] = "NYJ"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Las Vegas Raiders"] = "LVR"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Oakland Raiders"] = "LVR"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Philadelphia Eagles"] = "PHI"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Pittsburgh Steelers"] = "PIT"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Seattle Seahawks"] = "SEA"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "San Francisco 49ers"] = "SF"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Tampa Bay Buccaneers"] = "TB"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Tennessee Titans"] = "TEN"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Washington Redskins"] = "WAS"
NFL_Data_After2000$team_home[NFL_Data_After2000$team_home == "Washington Football Team"] = "WAS"

#Replace Away Team column with Away Team ID 
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Arizona Cardinals"] = "ARI"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Atlanta Falcons"] = "ATL"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Baltimore Ravens"] = "BAL"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Buffalo Bills"] = "BUF"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Carolina Panthers"] = "CAR"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Chicago Bears"] = "CHI"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Cincinnati Bengals"] = "CIN"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Cleveland Browns"] = "CLE"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Dallas Cowboys"] = "DAL"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Denver Broncos"] = "DEN"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Detroit Lions"] = "DET"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Green Bay Packers"] = "GB"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Houston Texans"] = "HOU"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Indianapolis Colts"] = "IND"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Jacksonville Jaguars"] = "JAX"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Kansas City Chiefs"] = "KC"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Los Angeles Chargers"] = "LAC"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "San Diego Chargers"] = "LAC"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Los Angeles Rams"] = "LAR"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "St. Louis Rams"] = "LAR"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Miami Dolphins"] = "MIA"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Minnesota Vikings"] = "MIN"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "New England Patriots"] = "NE"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "New Orleans Saints"] = "NO"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "New York Giants"] = "NYG"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "New York Jets"] = "NYJ"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Las Vegas Raiders"] = "LVR"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Oakland Raiders"] = "LVR"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Philadelphia Eagles"] = "PHI"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Pittsburgh Steelers"] = "PIT"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Seattle Seahawks"] = "SEA"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "San Francisco 49ers"] = "SF"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Tampa Bay Buccaneers"] = "TB"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Tennessee Titans"] = "TEN"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Washington Redskins"] = "WAS"
NFL_Data_After2000$team_away[NFL_Data_After2000$team_away == "Washington Football Team"] = "WAS"

#Create Column for if Home team won
NFL_Data_After2000 = NFL_Data_After2000 %>%
  mutate(Home_Team_Win = if_else(score_home>score_away,TRUE,FALSE))

#Create Column for Indoor Stadium or not
NFL_Data_After2000 = NFL_Data_After2000 %>%
  mutate(Indoor_Stadium = if_else(weather_detail == "DOME",TRUE,FALSE))

#Create Column for Indoor Stadium with roof open or not
NFL_Data_After2000 = NFL_Data_After2000 %>%
  mutate(Indoor_Stadium_Open = if_else(weather_detail == "DOME (Open Roof)",TRUE,FALSE))

```

##Exploratory Data Analysis
I will now utilize the created data frames in order to analyze the data and explore potential relations and correlations between different columns and data sets. The data frame consists of data including information about NFL games, stadiums, weather, betting, and more. There are many relations and scenarios that can be examined to look for potential correlation. If there are clear trends or correlations, these can be used to predict the outcomes of games and be a helpful tool for someone looking to bet on football games. There are many scenarios that I will be investigating, including:
-Favored Team Wins
-Home Team Wins when Favored
-Away Team Wins when Favored
-Spread is Covered given Favored Team Wins
-Losing Team Covers Spread
-Spread is a Push (Tie)
-Over Hits in Game Played in Dome
-Over Hits in Game Played in Dome (Open Roof)
-Over Hits in Game Played Outdoors
-Game is in Dome if Over Hits
-Game is in Dome (Open Roof) if Over Hits
-Game is Outdoors if Over hits
-Over Hits in Rain
-Over Hits in Snow
-Over Hits in Fog
-Over Hits in Rain and Fog
-Over Hits in Snow and Fog
-Over Hits in Clear Weather
-Favored Team Wins in Weeks 1-8
-Favored Team Wins in Weeks 9-17
-Favored Team Wins in Postseason
-Favored Team Covers Spread in Weeks 1-8
-Favored Team Covers Spread in Weeks 9-17
-Favored Team Covers Spread in Postseason
-Over Hits in Weeks 1-8
-Over Hits in Weeks 9-17
-Over Hits in Postseason
-Favorite Covers Spread in Rain
-Favorite Covers Spread in Snow
-Favorite Covers Spread in Fog
-Favorite Covers Spread in Rain and Fog
-Favorite Covers Spread in Snow and Fog
-Favorite Covers Spread in Clear Weather
-Favorite Covers Spread and Over Hits in Same Game



In this section, I will be exploring how often the team that is favored in the game wins. 
I will expand on this further by exploring how often the home team wins when favored versus how often the away team wins when favored.
```{r, echo=TRUE, eval=TRUE}
#How Often Favored Team wins
HomeTeam_FavoritedandWin = subset(NFL_Data_After2000, team_home == team_favorite_id & Home_Team_Win)
AwayTeam_FavoritedandWin = subset(NFL_Data_After2000, team_away == team_favorite_id & !Home_Team_Win)
Favorite_Wins = rbind(HomeTeam_FavoritedandWin, AwayTeam_FavoritedandWin)
nrow(Favorite_Wins)/nrow(NFL_Data_After2000)

#Create pie chart showing frequency of Home team winning when favored
labels = c('Favorite Wins','Underdog Wins')
values = c(nrow(Favorite_Wins)/nrow(NFL_Data_After2000),
           1-nrow(Favorite_Wins)/nrow(NFL_Data_After2000))

FigX = plot_ly(type='pie', labels=labels, values=values, title= "Favorite Team Wins/Losses", 
               textinfo='label+percent',
               insidetextorientation='radial')
FigX

#How Often Home Team Wins when favored
HomeTeam_Favored = subset(NFL_Data_After2000, team_home == team_favorite_id)
nrow(HomeTeam_FavoritedandWin)/nrow(HomeTeam_Favored)

#How Often Away Team wins when favored
AwayTeam_Favored = subset(NFL_Data_After2000, team_away == team_favorite_id)
nrow(AwayTeam_FavoritedandWin)/nrow(AwayTeam_Favored)

#Create pie chart showing frequency of Home team winning when favored
labels = c('Favorite at Home Wins','Favorite at Home Loses')
values = c(nrow(HomeTeam_FavoritedandWin)/nrow(HomeTeam_Favored),
           1-nrow(HomeTeam_FavoritedandWin)/nrow(HomeTeam_Favored))

Fig1 = plot_ly(type='pie', labels=labels, values=values, title= "Home Favorite Wins/Losses", 
               textinfo='label+percent',
               insidetextorientation='radial')
Fig1

#Create pie chart showing frequency of Away team winning when favored
labels = c('Away Favorite Wins','Away Favorite Loses')
values = c(nrow(AwayTeam_FavoritedandWin)/nrow(AwayTeam_Favored),
           1-nrow(AwayTeam_FavoritedandWin)/nrow(AwayTeam_Favored))

Fig2 = plot_ly(type='pie', labels=labels, values=values, title= "Away Favorite Wins/Losses", 
               textinfo='label+percent',
               insidetextorientation='radial')
Fig2
```

My next conjecture will be about how often the spread is covered when the favored team wins. I will also investigate how often the losing team is within the spread so they would win the bet.

```{r, echo=TRUE, eval=TRUE}
#How often the spread is covered when favored team wins
HomeFavoriteWins_andCovers = subset(HomeTeam_FavoritedandWin, score_home - score_away > abs(spread_favorite))
AwayFavoriteWins_andCovers = subset(AwayTeam_FavoritedandWin, score_away - score_home > abs(spread_favorite))
FavoriteWins_andCovers = rbind(HomeFavoriteWins_andCovers, AwayFavoriteWins_andCovers)
nrow(FavoriteWins_andCovers)/nrow(Favorite_Wins)

#How often is spread covered by losing team
#Favorites losing could never cover spread, only consider favored team wins
HomeFavoriteWins_LoserCovers = subset(HomeTeam_FavoritedandWin, score_home - score_away < abs(spread_favorite))
AwayFavoriteWins_LoserCovers = subset(AwayTeam_FavoritedandWin, score_away - score_home < abs(spread_favorite))
FavoriteWins_LoserCovers = rbind(HomeFavoriteWins_LoserCovers, AwayFavoriteWins_LoserCovers)
nrow(FavoriteWins_LoserCovers)/nrow(Favorite_Wins)

#How often spread is a push (causes tie)
HomeFavoriteWins_Push = subset(HomeTeam_FavoritedandWin, score_home - score_away == abs(spread_favorite))
AwayFavoriteWins_Push = subset(AwayTeam_FavoritedandWin, score_away - score_home == abs(spread_favorite))
FavoriteWins_Push = rbind(HomeFavoriteWins_Push, AwayFavoriteWins_Push)
nrow(FavoriteWins_Push)/nrow(Favorite_Wins)

#Create Pie chart showing frequency of spread being covered in favored team win
labels = c('Favorite Covers','Loser Covers','Push')
values = c(nrow(FavoriteWins_andCovers)/nrow(Favorite_Wins),
           nrow(FavoriteWins_LoserCovers)/nrow(Favorite_Wins),
           nrow(FavoriteWins_Push)/nrow(Favorite_Wins))

Fig3 = plot_ly(type='pie', labels=labels, values=values, title= "Results with Spread when Favorite Wins",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig3

```

My next conjecture will be exploring the frequency of the over being hit in games played indoors, indoors with the roof open, and outdoors. I will show the percentage of games played in a Dome that result in the over being hit, the percentage of games played in an open roof dome that result in the over being hit, and the percentage of games played outdoors that result in the over being hit

```{r, echo=TRUE, eval=TRUE}
#How often over is hit in games played in Dome
DomeGames = subset(NFL_Data_After2000, Indoor_Stadium == TRUE)
OverHit_Dome = subset(DomeGames, Total_Points > over_under_line)
nrow(OverHit_Dome)/nrow(DomeGames)

#Create Pie chart showing frequency of over being hit when game is played in a Dome
labels = c('Over Hit','Over not Hit')
values = c(nrow(OverHit_Dome)/nrow(DomeGames),
           1-nrow(OverHit_Dome)/nrow(DomeGames))

Fig4 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Over is Hit when Playing in Dome", 
               textinfo='label+percent',
               insidetextorientation='radial')
Fig4

#How often over is hit in games Played in Dome (Open Roof)
OpenRoofDome_Games = subset(NFL_Data_After2000, Indoor_Stadium_Open == TRUE)
OverHit_OpenRoofDome = subset(OpenRoofDome_Games, Total_Points > over_under_line)
nrow(OverHit_OpenRoofDome)/nrow(OpenRoofDome_Games)

#Create Pie chart showing frequency of over being hit when game is played in a Dome (Open Roof)
labels = c('Over Hit','Over not Hit')
values = c(nrow(OverHit_OpenRoofDome)/nrow(OpenRoofDome_Games),
           1-nrow(OverHit_OpenRoofDome)/nrow(OpenRoofDome_Games))

Fig5 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Over is Hit when Playing in Dome (Open Roof)",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig5

#How often over is hit in games played outdoors
OutdoorGames = subset(NFL_Data_After2000, Indoor_Stadium == FALSE & Indoor_Stadium_Open == FALSE)
OverHit_Outdoor = subset(OutdoorGames, Total_Points > over_under_line)
nrow(OverHit_Outdoor)/nrow(OutdoorGames)

#Create Pie chart showing frequency of over being hit when game is played Outdoors
labels = c('Over Hit','Over not Hit')
values = c(nrow(OverHit_Outdoor)/nrow(OutdoorGames),
           1-nrow(OverHit_Outdoor)/nrow(OutdoorGames))

Fig6 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Over is Hit when Playing Outdoors",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig6

```

I will next look at what percentage of the time the over is hit is it indoors, outdoors, or indoors with an open roof

```{r, echo=TRUE, eval=TRUE}
#Create Subset of all games where over is hit
Over_Hit = subset(NFL_Data_After2000, Total_Points > over_under_line)
nrow(Over_Hit)/nrow(NFL_Data_After2000)

#Percentage of time game is in Dome if over is hit
nrow(OverHit_Dome)/nrow(Over_Hit)

#Percentage of time game is in Dome (Roof Open) if over  is hit
nrow(OverHit_OpenRoofDome)/nrow(Over_Hit)

#Percentage of time game is Outdoors if over is hit
nrow(OverHit_Outdoor)/nrow(Over_Hit)

#Create Pie Chart showing percentage of the time the game is at each stadium type given the over hits
labels = c('Dome','Dome (Open Roof)','Outdoor')
values = c(nrow(OverHit_Dome)/nrow(Over_Hit),
           nrow(OverHit_OpenRoofDome)/nrow(Over_Hit),
           nrow(OverHit_Outdoor)/nrow(Over_Hit))

Fig7 = plot_ly(type='pie', labels=labels, values=values, title= "Stadium Type Likelihood if Over Hits",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig7


```

The next conjecture I will investigate is the impact of precipitation and fog on the Over/Under hitting

```{r, echo=TRUE, eval=TRUE}
#Create Column for if it is raining
NFL_Data_After2000 = NFL_Data_After2000 %>%
  mutate(Rain = if_else(weather_detail == "Rain",TRUE,FALSE))

#Create Column for if it is snowing
NFL_Data_After2000 = NFL_Data_After2000 %>%
  mutate(Snow = if_else(weather_detail == "Snow",TRUE,FALSE))

##Create Column for if it is Foggy
NFL_Data_After2000 = NFL_Data_After2000 %>%
  mutate(Fog = if_else(weather_detail == "Fog",TRUE,FALSE))

#Create Column for if it is Rain and Fog
NFL_Data_After2000 = NFL_Data_After2000 %>%
  mutate(Rain_Fog = if_else(weather_detail == "Rain | Fog",TRUE,FALSE))

#Create Column for if it is Snow and Fog
NFL_Data_After2000 = NFL_Data_After2000 %>%
  mutate(Snow_Fog = if_else(weather_detail == "Snow | Fog",TRUE,FALSE))

#Create Column for if it is Snow and Freezing Rain
NFL_Data_After2000 = NFL_Data_After2000 %>%
  mutate(Snow_FreezingRain = if_else(weather_detail == "Snow | Freezing Rain",TRUE,FALSE))

#Frequency of Over hitting if Raining
RainGames = subset(NFL_Data_After2000, Rain == TRUE)
OverHit_Rain = subset(RainGames, Total_Points > over_under_line)
nrow(OverHit_Rain)/nrow(RainGames)

#Frequency of Over hitting if Snowing
SnowGames = subset(NFL_Data_After2000, Snow == TRUE)
OverHit_Snow = subset(SnowGames, Total_Points > over_under_line)
nrow(OverHit_Snow)/nrow(SnowGames)

#Frequency of Over hitting if Foggy
FogGames = subset(NFL_Data_After2000, Fog == TRUE)
OverHit_Fog = subset(FogGames, Total_Points > over_under_line)
nrow(OverHit_Fog)/nrow(FogGames)

#Frequency of Over hitting if Raining and Foggy
RainFog_Games = subset(NFL_Data_After2000, Rain_Fog == TRUE)
OverHit_RainFog = subset(RainFog_Games, Total_Points > over_under_line)
nrow(OverHit_RainFog)/nrow(RainFog_Games)

#Frequency of Over hitting if Snowing and Foggy
SnowFog_Games = subset(NFL_Data_After2000, Snow_Fog == TRUE)
OverHit_SnowFog = subset(SnowFog_Games, Total_Points > over_under_line)
nrow(OverHit_SnowFog)/nrow(SnowFog_Games)

#Frequency of Over hitting if Clear weather
ClearGames = subset(NFL_Data_After2000, Rain_Fog == FALSE & Snow_Fog == FALSE & Snow == FALSE & Rain == FALSE & Fog == FALSE & Snow_FreezingRain == FALSE )
OverHit_Clear = subset(ClearGames, Total_Points > over_under_line)
nrow(OverHit_Clear)/nrow(ClearGames)

#Create Pie chart for frequency of over hitting in Rain Games
labels = c('Over Hit','Over not Hit')
values = c(nrow(OverHit_Rain)/nrow(RainGames),
           1-nrow(OverHit_Rain)/nrow(RainGames))

Fig8 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Over is Hit when Raining",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig8

#Create Pie chart for frequency of over hitting in Snow Games
labels = c('Over Hit','Over not Hit')
values = c(nrow(OverHit_Snow)/nrow(SnowGames),
           1-nrow(OverHit_Snow)/nrow(SnowGames))

Fig9 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Over is Hit when Snowing",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig9

#Create Pie chart for frequency of over hitting in Fog Games
labels = c('Over Hit','Over not Hit')
values = c(nrow(OverHit_Fog)/nrow(FogGames),
           1-nrow(OverHit_Fog)/nrow(FogGames))

Fig10 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Over is Hit when Foggy",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig10

#Create Pie chart for frequency of over hitting in Games where it is raining and foggy
labels = c('Over Hit','Over not Hit')
values = c(nrow(OverHit_RainFog)/nrow(RainFog_Games),
           1-nrow(OverHit_RainFog)/nrow(RainFog_Games))

Fig11 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Over is Hit when Raining and Foggy",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig11

#Create Pie chart for frequency of over hitting in Games where it is snowing and foggy
labels = c('Over Hit','Over not Hit')
values = c(nrow(OverHit_SnowFog)/nrow(SnowFog_Games),
           1-nrow(OverHit_SnowFog)/nrow(SnowFog_Games))

Fig12 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Over is Hit when Snowing and Foggy",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig12

#Create Pie chart for frequency of over hitting in Games where it is Clear
labels = c('Over Hit','Over not Hit')
values = c(nrow(OverHit_Clear)/nrow(ClearGames),
           1-nrow(OverHit_Clear)/nrow(ClearGames))

Fig13 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Over is Hit when Clear",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig13

```

The next relation I will be exploring is whether the favorite team is more likely to win in the first half or second half of the season. I will find the percentage of the time the favored team won in weeks 1-8 and the percentage of the time the favored team won in weeks 9-17. I will then look at percentage of the time the favored team won in the postseason.


```{r, echo=TRUE, eval=TRUE}
#Create Subset of just regular season games and Subset of postseason games
RegularSeason_Games = subset(NFL_Data_After2000, schedule_playoff == FALSE)
PostSeason_Games = subset(NFL_Data_After2000, schedule_playoff == TRUE)

#Frequency of Favored Team winning in weeks 1-8
FirstHalf_Games = subset(RegularSeason_Games, schedule_week == "1" | schedule_week == "2" |  schedule_week == "3" |  schedule_week == "4" |  schedule_week == "5" |  schedule_week == "6" |  schedule_week == "7" |  schedule_week == "8")
HomeTeam_FavoredWin_First = subset(FirstHalf_Games, team_home == team_favorite_id & Home_Team_Win)
AwayTeam_FavoredWin_First = subset(FirstHalf_Games, team_away == team_favorite_id & !Home_Team_Win)
Favorite_Wins_First = rbind(HomeTeam_FavoredWin_First, AwayTeam_FavoredWin_First)
nrow(Favorite_Wins_First)/nrow(FirstHalf_Games)

#Frequency of Favored Team winning in weeks 9-17
SecondHalf_Games = subset(RegularSeason_Games, schedule_week == "9" | schedule_week == "10" |  schedule_week == "11" |  schedule_week == "12" |  schedule_week == "13" |  schedule_week == "14" |  schedule_week == "15" |  schedule_week == "16" | schedule_week == "17")
HomeTeam_FavoredWin_Second = subset(SecondHalf_Games, team_home == team_favorite_id & Home_Team_Win)
AwayTeam_FavoredWin_Second = subset(SecondHalf_Games, team_away == team_favorite_id & !Home_Team_Win)
Favorite_Wins_Second = rbind(HomeTeam_FavoredWin_Second, AwayTeam_FavoredWin_Second)
nrow(Favorite_Wins_Second)/nrow(SecondHalf_Games)

#Frequency of Favored Team winning in PostSeason
HomeTeam_FavoredWin_PostSeason = subset(PostSeason_Games, team_home == team_favorite_id & Home_Team_Win)
AwayTeam_FavoredWin_PostSeason = subset(PostSeason_Games, team_away == team_favorite_id & !Home_Team_Win)
Favorite_Wins_PostSeason = rbind(HomeTeam_FavoredWin_PostSeason, AwayTeam_FavoredWin_PostSeason)
nrow(Favorite_Wins_PostSeason)/nrow(PostSeason_Games)

#Create Pie chart for frequency of favored team winning in weeks 1-8
labels = c('Favorite Win','Favorite Loss')
values = c(nrow(Favorite_Wins_First)/nrow(FirstHalf_Games),
           1-nrow(Favorite_Wins_First)/nrow(FirstHalf_Games))

Fig14 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Favorite Wins in Weeks 1-8",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig14

#Create Pie chart for frequency of favored team winning in weeks 9-17
labels = c('Favorite Win','Favorite Loss')
values = c(nrow(Favorite_Wins_Second)/nrow(SecondHalf_Games),
           1-nrow(Favorite_Wins_Second)/nrow(SecondHalf_Games))

Fig15 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Favorite Wins in Weeks 9-17",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig15

#Create Pie chart for frequency of favored team winning in PostSeason
labels = c('Favorite Win','Favorite Loss')
values = c(nrow(Favorite_Wins_PostSeason)/nrow(PostSeason_Games),
           1-nrow(Favorite_Wins_PostSeason)/nrow(PostSeason_Games))

Fig16 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Favorite Wins in Postseason",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig16
```

The next conjecture I will make is how often the favored team covers the spread in different parts of the season. I will once again be investigating weeks 1-8, weeks 9-17, and the postseason.

```{r, echo=TRUE, eval=TRUE}
#Frequency of favored team covering spread in weeks 1-8
HomeTeam_FavoredWinCover_First = subset(HomeTeam_FavoredWin_First, score_home - score_away > abs(spread_favorite))
AwayTeam_FavoredWinCover_First = subset(AwayTeam_FavoredWin_First, score_away - score_home > abs(spread_favorite))
Favorite_WinsCovers_First = rbind(HomeTeam_FavoredWinCover_First, AwayTeam_FavoredWinCover_First)
nrow(Favorite_WinsCovers_First)/nrow(FirstHalf_Games)

#Frequency of favored team covering spread in weeks 9-17
HomeTeam_FavoredWinCover_Second = subset(HomeTeam_FavoredWin_Second, score_home - score_away > abs(spread_favorite))
AwayTeam_FavoredWinCover_Second = subset(AwayTeam_FavoredWin_Second, score_away - score_home > abs(spread_favorite))
Favorite_WinsCovers_Second = rbind(HomeTeam_FavoredWinCover_Second, AwayTeam_FavoredWinCover_Second)
nrow(Favorite_WinsCovers_Second)/nrow(SecondHalf_Games)

#Frequency of favored team covering spread in Postseason
HomeTeam_FavoredWinCover_PostSeason = subset(HomeTeam_FavoredWin_PostSeason, score_home - score_away > abs(spread_favorite))
AwayTeam_FavoredWinCover_PostSeason = subset(AwayTeam_FavoredWin_PostSeason, score_away - score_home > abs(spread_favorite))
Favorite_WinsCovers_PostSeason = rbind(HomeTeam_FavoredWinCover_PostSeason, AwayTeam_FavoredWinCover_PostSeason)
nrow(Favorite_WinsCovers_PostSeason)/nrow(PostSeason_Games)

#Create Pie chart for frequency of favored team covering in weeks 1-8
labels = c('Favorite Covers','Favorite Does Not Cover')
values = c(nrow(Favorite_WinsCovers_First)/nrow(FirstHalf_Games),
           1-nrow(Favorite_WinsCovers_First)/nrow(FirstHalf_Games))

Fig17 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Favorite Covers in Weeks 1-8",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig17

#Create Pie chart for frequency of favored team covering in weeks 9-17
labels = c('Favorite Covers','Favorite Does Not Cover')
values = c(nrow(Favorite_WinsCovers_Second)/nrow(SecondHalf_Games),
           1-nrow(Favorite_WinsCovers_Second)/nrow(SecondHalf_Games))

Fig18 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Favorite Covers in Weeks 9-17",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig18

#Create Pie chart for frequency of favored team covering in Postseason
labels = c('Favorite Covers','Favorite Does Not Cover')
values = c(nrow(Favorite_WinsCovers_PostSeason)/nrow(PostSeason_Games),
           1-nrow(Favorite_WinsCovers_PostSeason)/nrow(PostSeason_Games))

Fig19 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Favorite Covers in Postseason",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig19
```

I will next investigate the frequency with which the over hits in different parts of the season. I will look at weeks 1-8, weeks 9-17, and the postseason.

```{r, echo=TRUE, eval=TRUE}
#Frequency of Over hitting in Weeks 1-8
OverHit_First = subset(FirstHalf_Games, Total_Points > over_under_line)
nrow(OverHit_First)/nrow(FirstHalf_Games)

#Frequency of Over hitting in Weeks 9-17
OverHit_Second = subset(SecondHalf_Games, Total_Points > over_under_line)
nrow(OverHit_Second)/nrow(SecondHalf_Games)

#Frequency of Over hitting in Postseason
OverHit_PostSeason = subset(PostSeason_Games, Total_Points > over_under_line)
nrow(OverHit_PostSeason)/nrow(PostSeason_Games)

#Create Pie chart for frequency of over hitting in weeks 1-8
labels = c('Over Hit','Over not Hit')
values = c(nrow(OverHit_First)/nrow(FirstHalf_Games),
           1-nrow(OverHit_First)/nrow(FirstHalf_Games))

Fig20 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Over is Hit Weeks 1-8",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig20

#Create Pie chart for frequency of over hitting in weeks 9-17
labels = c('Over Hit','Over not Hit')
values = c(nrow(OverHit_Second)/nrow(SecondHalf_Games),
           1-nrow(OverHit_Second)/nrow(SecondHalf_Games))

Fig21 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Over is Hit Weeks 9-17",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig21

#Create Pie chart for frequency of over hitting in Postseason
labels = c('Over Hit','Over not Hit')
values = c(nrow(OverHit_PostSeason)/nrow(PostSeason_Games),
           1-nrow(OverHit_PostSeason)/nrow(PostSeason_Games))

Fig22 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency Over is Hit in Postseason",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig22
```

Another conjecture I make is how does the do different weather types impact the closeness of games? To investigate this I will be evaluating the frequency with which the favorite covers the spread in each whether type with sufficient data.

```{r, echo=TRUE, eval=TRUE}
#Frequency of Favorite covering spread if Raining
HomeFavoriteWin_Rain = subset(RainGames, team_home == team_favorite_id & Home_Team_Win)
AwayFavoriteWin_Rain = subset(RainGames, team_away == team_favorite_id & !Home_Team_Win)
FavoriteWin_Rain = rbind(HomeFavoriteWin_Rain, AwayFavoriteWin_Rain)
HomeFavoriteCover_Rain = subset(HomeFavoriteWin_Rain, score_home - score_away > abs(spread_favorite))
AwayFavoriteCover_Rain = subset(AwayFavoriteWin_Rain, score_away - score_home > abs(spread_favorite))
FavoriteWinCover_Rain = rbind(HomeFavoriteCover_Rain, AwayFavoriteCover_Rain)
nrow(FavoriteWinCover_Rain)/nrow(RainGames)

#Frequency of Favorite covering spread if Snowing
HomeFavoriteWin_Snow = subset(SnowGames, team_home == team_favorite_id & Home_Team_Win)
AwayFavoriteWin_Snow = subset(SnowGames, team_away == team_favorite_id & !Home_Team_Win)
FavoriteWin_Snow = rbind(HomeFavoriteWin_Snow, AwayFavoriteWin_Snow)
HomeFavoriteCover_Snow = subset(HomeFavoriteWin_Snow, score_home - score_away > abs(spread_favorite))
AwayFavoriteCover_Snow = subset(AwayFavoriteWin_Snow, score_away - score_home > abs(spread_favorite))
FavoriteWinCover_Snow = rbind(HomeFavoriteCover_Snow, AwayFavoriteCover_Snow)
nrow(FavoriteWinCover_Snow)/nrow(SnowGames)

#Frequency of Favorite covering spread if Foggy
HomeFavoriteWin_Fog = subset(FogGames, team_home == team_favorite_id & Home_Team_Win)
AwayFavoriteWin_Fog = subset(FogGames, team_away == team_favorite_id & !Home_Team_Win)
FavoriteWin_Fog = rbind(HomeFavoriteWin_Fog, AwayFavoriteWin_Fog)
HomeFavoriteCover_Fog = subset(HomeFavoriteWin_Fog, score_home - score_away > abs(spread_favorite))
AwayFavoriteCover_Fog = subset(AwayFavoriteWin_Fog, score_away - score_home > abs(spread_favorite))
FavoriteWinCover_Fog = rbind(HomeFavoriteCover_Fog, AwayFavoriteCover_Fog)
nrow(FavoriteWinCover_Fog)/nrow(FogGames)

#Frequency of Favorite covering spread if Raining and Foggy
HomeFavoriteWin_RainFog = subset(RainFog_Games, team_home == team_favorite_id & Home_Team_Win)
AwayFavoriteWin_RainFog = subset(RainFog_Games, team_away == team_favorite_id & !Home_Team_Win)
FavoriteWin_RainFog = rbind(HomeFavoriteWin_RainFog, AwayFavoriteWin_RainFog)
HomeFavoriteCover_RainFog = subset(HomeFavoriteWin_RainFog, score_home - score_away > abs(spread_favorite))
AwayFavoriteCover_RainFog = subset(AwayFavoriteWin_RainFog, score_away - score_home > abs(spread_favorite))
FavoriteWinCover_RainFog = rbind(HomeFavoriteCover_RainFog, AwayFavoriteCover_RainFog)
nrow(FavoriteWinCover_RainFog)/nrow(RainFog_Games)

#Frequency of Favorite covering spread if Snowing and Foggy
HomeFavoriteWin_SnowFog = subset(SnowFog_Games, team_home == team_favorite_id & Home_Team_Win)
AwayFavoriteWin_SnowFog = subset(SnowFog_Games, team_away == team_favorite_id & !Home_Team_Win)
FavoriteWin_SnowFog = rbind(HomeFavoriteWin_SnowFog, AwayFavoriteWin_SnowFog)
HomeFavoriteCover_SnowFog = subset(HomeFavoriteWin_SnowFog, score_home - score_away > abs(spread_favorite))
AwayFavoriteCover_SnowFog = subset(AwayFavoriteWin_SnowFog, score_away - score_home > abs(spread_favorite))
FavoriteWinCover_SnowFog = rbind(HomeFavoriteCover_SnowFog, AwayFavoriteCover_SnowFog)
nrow(FavoriteWinCover_SnowFog)/nrow(SnowFog_Games)

#Frequency of Favorite covering spread if Clear
HomeFavoriteWin_Clear = subset(ClearGames, team_home == team_favorite_id & Home_Team_Win)
AwayFavoriteWin_Clear = subset(ClearGames, team_away == team_favorite_id & !Home_Team_Win)
FavoriteWin_Clear = rbind(HomeFavoriteWin_Clear, AwayFavoriteWin_Clear)
HomeFavoriteCover_Clear = subset(HomeFavoriteWin_Clear, score_home - score_away > abs(spread_favorite))
AwayFavoriteCover_Clear = subset(AwayFavoriteWin_Clear, score_away - score_home > abs(spread_favorite))
FavoriteWinCover_Clear = rbind(HomeFavoriteCover_Clear, AwayFavoriteCover_Clear)
nrow(FavoriteWinCover_Clear)/nrow(ClearGames)

#Create Pie chart for frequency of Favorite covering spread in Rain Games
labels = c('Favorite Covered','Favorite Did Not Cover')
values = c(nrow(FavoriteWinCover_Rain)/nrow(RainGames),
           1-nrow(FavoriteWinCover_Rain)/nrow(RainGames))

Fig23 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency of Favorite Covering Spread when Raining",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig23

#Create Pie chart for frequency of Favorite covering spread in Snow Games
labels = c('Favorite Covered','Favorite Did Not Cover')
values = c(nrow(FavoriteWinCover_Snow)/nrow(SnowGames),
           1-nrow(FavoriteWinCover_Snow)/nrow(SnowGames))

Fig24 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency of Favorite Covering Spread when Snowing",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig24

#Create Pie chart for frequency of Favorite covering spread in Foggy Games
labels = c('Favorite Covered','Favorite Did Not Cover')
values = c(nrow(FavoriteWinCover_Fog)/nrow(FogGames),
           1-nrow(FavoriteWinCover_Fog)/nrow(FogGames))

Fig25 = plot_ly(type='pie', labels=labels, values=values, title= "Frequency of Favorite Covering Spread when Foggy",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig25

#Create Pie chart for frequency of Favorite covering spread in Games where it is raining and foggy
labels = c('Favorite Covered','Favorite Did Not Cover')
values = c(nrow(FavoriteWinCover_RainFog)/nrow(RainFog_Games),
           1-nrow(FavoriteWinCover_RainFog)/nrow(RainFog_Games))

Fig26 = plot_ly(type='pie', labels=labels, values=values, 
               title= "Frequency of Favorite Covering Spread when Raining and Foggy",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig26

#Create Pie chart for frequency of Favorite covering spread in Games where it is snowing and foggy
labels = c('Favorite Covered','Favorite Did Not Cover')
values = c(nrow(FavoriteWinCover_SnowFog)/nrow(SnowFog_Games),
           1-nrow(FavoriteWinCover_SnowFog)/nrow(SnowFog_Games))

Fig27 = plot_ly(type='pie', labels=labels, values=values, 
               title= "Frequency of Favorite Covering Spread when Snowing and Foggy",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig27

#Create Pie chart for frequency of Favorite covering spread in Games where it is clear
labels = c('Favorite Covered','Favorite Did Not Cover')
values = c(nrow(FavoriteWinCover_Clear)/nrow(ClearGames),
           1-nrow(FavoriteWinCover_Clear)/nrow(ClearGames))

Fig28 = plot_ly(type='pie', labels=labels, values=values, 
               title= "Frequency of Favorite Covering Spread when Clear",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig28

```

Next, I am going to investigate how frequently the favored team covers the spread and the over hits in the same game.

```{r, echo=TRUE, eval=TRUE}
#Over Hitting and Favorite covers spread
HomeFavoriteWin = subset(NFL_Data_After2000, team_home == team_favorite_id & Home_Team_Win)
AwayFavoriteWin = subset(NFL_Data_After2000, team_away == team_favorite_id & !Home_Team_Win)
FavoriteWin = rbind(HomeFavoriteWin, AwayFavoriteWin)
HomeFavoriteCover = subset(HomeFavoriteWin, score_home - score_away > abs(spread_favorite))
AwayFavoriteCover = subset(AwayFavoriteWin, score_away - score_home > abs(spread_favorite))
FavoriteWinCover = rbind(HomeFavoriteCover, AwayFavoriteCover)
FavoriteWinCover_OverHits = subset(FavoriteWinCover, Total_Points > over_under_line)
nrow(FavoriteWinCover_OverHits)/nrow(NFL_Data_After2000)

#Create Pie chart for frequency of Favorite covering spread and Over hitting in same game
labels = c('Favorite Covered and Over Hit','Did not Happen')
values = c(nrow(FavoriteWinCover_OverHits)/nrow(NFL_Data_After2000),
           1-nrow(FavoriteWinCover_OverHits)/nrow(NFL_Data_After2000))

Fig29 = plot_ly(type='pie', labels=labels, values=values, 
               title= "Frequency of Favorite Covering Spread and Over Hitting in Same Game",
               textinfo='label+percent',
               insidetextorientation='radial')
Fig29

```

```{r, echo=TRUE, eval=TRUE}
#Create Variables for Scenarios investigated and Pie Chart Figures
Scenario = c("Favored Team Wins","Home Team Wins when Favored", "Away Team Wins when Favored", "Spread is Covered given Favored Team Wins", "Losing Team Covers Spread", "Spread is a Push (Tie)", "Over Hits in Game Played in Dome", "Over Hits in Game Played in Dome (Open Roof)", "Over Hits in Game Played Outdoors", "Game is in Dome if Over Hits", "Game is in Dome (Open Roof) if Over Hits", "Game is Outdoors if Over hits", "Over Hits in Rain", "Over Hits in Snow", "Over Hits in Fog", "Over Hits in Rain and Fog", "Over Hits in Snow and Fog", "Over Hits in Clear Weather", "Favored Team Wins in Weeks 1-8", "Favored Team Wins in Weeks 9-17", "Favored Team Wins in Postseason", "Favored Team Covers Spread in Weeks 1-8", "Favored Team Covers Spread in Weeks 9-17", "Favored Team Covers Spread in Postseason", "Over Hits in Weeks 1-8", "Over Hits in Weeks 9-17", "Over Hits in Postseason", "Favorite Covers Spread in Rain", "Favorite Covers Spread in Snow", "Favorite Covers Spread in Fog", "Favorite Covers Spread in Rain and Fog", "Favorite Covers Spread in Snow and Fog", "Favorite Covers Spread in Clear Weather", "Favorite Covers Spread and Over Hits in Same Game")

Figure = c("FigX", "Fig1", "Fig2", "Fig3", "Fig3", "Fig3", "Fig4", "Fig5", "Fig6", "Fig7", "Fig7", "Fig7", "Fig8", "Fig9", "Fig10", "Fig11", "Fig12", "Fig13", "Fig14", "Fig15", "Fig16", "Fig17", "Fig18", "Fig19", "Fig20", "Fig21", "Fig22", "Fig23", "Fig24", "Fig25", "Fig26", "Fig27", "Fig28", "Fig29")

#Create Dataframe of Scenarios and corresponding Pie Chart Figures
Scenarios_Charts = data.frame(Scenario, Figure)
Scenarios_Charts

#Assign Character to Figure
assign("FigX",FigX)
assign("Fig1",Fig1)
assign("Fig2",Fig2)
assign("Fig3",Fig3)
assign("Fig4",Fig4)
assign("Fig5",Fig5)
assign("Fig6",Fig6)
assign("Fig7",Fig7)
assign("Fig8",Fig8)
assign("Fig9",Fig9)
assign("Fig10",Fig10)
assign("Fig11",Fig11)
assign("Fig12",Fig12)
assign("Fig13",Fig13)
assign("Fig14",Fig14)
assign("Fig15",Fig15)
assign("Fig16",Fig16)
assign("Fig17",Fig17)
assign("Fig18",Fig18)
assign("Fig19",Fig19)
assign("Fig20",Fig20)
assign("Fig21",Fig21)
assign("Fig22",Fig22)
assign("Fig23",Fig23)
assign("Fig24",Fig24)
assign("Fig25",Fig25)
assign("Fig26",Fig26)
assign("Fig27",Fig27)
assign("Fig28",Fig28)
assign("Fig29",Fig29)
```

##Set up Shiny R

```{r, echo=TRUE, eval=TRUE}
library(shiny)
library(dplyr)

ui = fluidPage(
  #Navigation Bar
  navbarPage("NFL Data and Betting",
          #Tab One
          tabPanel("App",
             #Header Panel
             titlePanel(title = h3("NFL Games and Betting Results and Trends For Games after Year 2000", align="center")),
             titlePanel(title = h5("Created By Louis Duh", align="center")),
             sidebarLayout(
             #Sidebar Panel
             sidebarPanel(
             selectInput("Scenario", "Select a Scenario", Scenarios_Charts$Scenario, multiple=F, selected="Favored Team                            Wins")),
             #Main Panel
             mainPanel(
             plotlyOutput("Pie_Chart"))
            )
            ),
          #Tab Two
          tabPanel("About",
                tags$div("This site provides insight into performance trends in different areas of NFL games and Betting",
                         tags$br(), " ", tags$br(),
                        "This site was created as a deliverable for a Case Study in the course DS 501 at Worcester                                        Polytechnic Institute"),
                         tags$br(), " ", tags$br(),
                        "Further information and instructions on how to use this application can be found in the tutorial tab",
                         tags$br(), " ", tags$br(),
                        "The data that I collected for use in this case study and application came from a data set on Kaggle.com. The data set was titled 'NFL scores and betting data'. This data set included information such as the date and season year of the game, in what week of the season it took place, the names of the home and away teams and their scores, betting data such as the spread and over/under line, and stadium and weather information. This data set contained all games going back to the 1966 season, however betting data didn't start until 1978.",
                         tags$br(), " ", tags$br(),
                        "This topic is interesting to me because I love almost anything to do with the sport of football. I have played football for 14 years and have been constantly surrounded by the game. Sunday afternoons have been focused around watching as many NFL games as possible for years. One of the most interesting parts of the games to me is tracking and analyzing statistics. This goes beyond player statistics. I am quite interested in how different factors impact a team's chances to win a football game and the amount of scoring in the game. I am also fascinated by the concept of oddsmaking and Betting. Athough not a gambling myself, I find the concept and decision making of what to bet on to be incredibly interesting. I believe it is extra interesting when data and trends between scenarios and conditions can be analyzed to help choose bets.",
                         tags$br(), " ", tags$br(),
                        "I analyzed the data in a couple of ways. The first step was to clean the data so that it was suitable for use for my purposes and the observations I wanted to make. To do this, I made a subset data frame that restricted the data to only games that happened in the year 2000 or after. I chose this year as the cutoff because that is about when the modern passing era of the NFL began so it would make the model more true to today's NFL. I had to add some columns to the data frame in order to make analysis easier at later steps. Some of these columns were: If Home Team Won, Total Points, and individual columsn for each stadium type. I then changed the team names of the home and away team columns to be the two or three letter IDs so that they would match up with the existing colum 'favorite team id'. The main way I analyzed the data was to create subsets of the data frame that reflected the conditions that needed to be met for a given scenario. I then divided the number of items in that specific subset by the number in the collective subset for that condition set. I created pie charts for every scenario and analysis I completed using plotly.",
                         tags$br(), " ", tags$br(),
                        "There were many observations that I made based on the trends found from testing scenarios. I will name a few of them here. One observation is that the over tends to hit more frequently in games where the weather is not clear (Rain, Snow, fog, etc.). This is the opposite of what I thought would be true because bad weather is usually associated with less productive offense. I think this shows that the oddsmakers often overcompensate for the bad weatehr when making the betting lines for that game. This causes the over to be hit more frequently with less points scored than usual. Another observation is that favorites tend to win more often in Weeks 9-17 of the season than they do in weeks 1-8 and in the postseason. This could be because after the halfway point of the season, enough about the teams is known to begin making more accurate odds. Another big observation I make from the data is that Favored teams cover the spread far less in clear weather than they do in inclement weather such as rain, fog, or snow. In clear weather, the favorite covers about 46% of the time, while for all types of inclement weather, the favorite covers a percentage in the 50s. This could indicate that better teams play better and handle bad weather better than bad teams.",
                        tags$br(), " ", tags$br(),
                       "Code for this app can be found at: https://github.com/LDuh97/DS501-Case-Study-3",
                        tags$br(), " ", tags$br(),
                       "Author: Louis Duh,  WPI '21"
            ),
          #Tab Three
          tabPanel("Tutorial",
              tags$div("This app allows you to explore performance trends for NFL game stats and betting information",
                        tags$br(), " ", tags$br(),
                "How To Use:",
                        tags$br(),
                "Use dropdown box to select the scenario you are interested in exploring. There are 34 options to choose from and range from a simple 'How often Favored team wins' to How often favored team covers the spread in a variety of different weather types.",
                        tags$br(), " ", tags$br(),
                "Once you select the scenario you would like to consider, a pie chart will appear in the main panel of the screen that will show the frequency/likelihood of the scenario occurring versus not. The pie chart will include a legend that clearly identifies which color of the chart represents which outcome",
                        tags$br(), " ", tags$br(),
                "Explore all of the different scenarios, especially the ones you would have never thought to consider. These trends and data can be used as a predictive betting model if you are looking to get into NFL betting. Alternatively, If you are looking to be the envy of your friends and be able to predict the various outcomes of NFL games and their betting lines, this app will work great for you.",
                        tags$br(), " ", tags$br(),
                "Enjoy exploring the trends and please bet responsibly")
          
            )))


server = function(input,output){
  rowChosen = reactive({which(Scenarios_Charts$Scenario == input$Scenario)})
  output$Pie_Chart = renderPlotly(get(Scenarios_Charts[rowChosen(), 2]))
}

shinyApp(ui=ui, server=server)




```